The five pillars of computational reproducibility: Bioinformatics and beyond

### 2023-05-30 manuscript/5pillars.Rmd
more refs

A number of guides recommending enhanced computational reproducibility practices have emerged
[ @Tan2010-is; @Peng2011-ia; @Sandve2013-lk; @Piccolo2016-ji; @Lewis2016-qh;
@Wilson2017-uq ; @Akalin2018-eg ; @Gruning2018-sg; @Brito2020-ld; @Peikert2021-eq ; @Rodrigues2023-fw].
This article, inspired by the tutorial by Peikert and BrandMaier @Peikert2021-eq, attempts to unify
these concepts into five key principles called the five pillars of reproducible computational
research (**Figure 1**).
The intended audience is bioinformatics data analysts (not tool developers), however the
principles described here could equally apply to other domains of study.

@@

Without data sharing, computational research is not reproducible nor auditable.
Lack of data sharing is one of the key reasons why research is irreproducible
[@Ioannidis2009-hv, @Archmiller2020-er ].
Data sharing is also one of the key features of "open science", which is characterised as science
that is collaborative, transparent, accessible and inclusive [@noauthor_2023-uq].
In addition to facilitating reproduction and auditing, sharing enables reuse in other contexts.
Data reuse increases efficiency, as it prevents redundant research expenditure and facilitates new
research ideas that were previously impossible [@Wilkinson2016-qq].
In terms of research rigour, inspecting raw data can uncover research integrity problems [@Miyakawa2020-yj],
In light of this, a research article without supporting data and code is much like a press release or
advertisement where the claims made cannot be verified as true or not.

@@

I am grateful to the individuals who provided feedback on an earlier draft of the concept including
Dr Aziz Khan (Stanford University),
Dr Sriganesh Srihari (QIMR Berghofer Medical Research Institute),
Dr Barbara Zdrazil (EMBL-EBI)

* Irene López Santiago @IreneLopSci

>>>>

A number of guides recommending enhanced computational reproducibility practices have emerged
[ @Tan2010-is; @Peng2011-ia; @Sandve2013-lk; @Piccolo2016-ji; @Lewis2016-qh;
@Wilson2017-uq ; @Akalin2018-eg ; @Gruning2018-sg; @Brito2020-ld; @Peikert2021-eq ; @Rodrigues2023-fw;
@DuPre2022-bk].
Inspired by these, this article attempts to unify the main principles into the
*five pillars of reproducible computational research* (**Figure 1**).
The intended audience is bioinformatics data analysts (not tool developers), however the
principles described here could equally apply to other domains of study.

@@

Without data sharing, computational research is not reproducible nor auditable.
Lack of data sharing is one of the key reasons why research is irreproducible
[@Ioannidis2009-hv, @Archmiller2020-er ].
Data sharing is also one of the key features of "open science", which is characterised as science
that is collaborative, transparent, accessible and inclusive [@noauthor_2023-uq].
In addition to facilitating reproduction and auditing, sharing enables reuse in other contexts.
Data reuse increases efficiency, as it prevents redundant research expenditure and facilitates new
research ideas that were previously impossible [@Wilkinson2016-qq].
In terms of research rigour, inspecting raw data can uncover research integrity problems [@Miyakawa2020-yj],
In light of this, a research article without supporting data and code is much like a press release or
advertisement where the claims made cannot be verified as true or not [@Donoho2010-yz, @DuPre2022-bk].

@@

I am grateful to the individuals who provided feedback on an earlier draft of the concept including
Dr Aziz Khan (Stanford University),
Dr Sriganesh Srihari (QIMR Berghofer Medical Research Institute),
Dr Barbara Zdrazil (EMBL-EBI) and Dr Irene López Santiago (EMBL-EBI).
